NLTK: The Natural Language Toolkit

### 2002-03-18 technical/acl2002/submission.tex
Added section on the chunking project, based on LSA report
svn/trunk@497

\documentclass[11pt]{article}
\usepackage{acl02sub,url}
\title{NLTK: The Natural Language Toolkit}
\author{
Edward Loper and Steven Bird\\
Department of Computer and Information Science and Linguistic Data Consortium\\
University of Pennsylvania, Philadelphia, PA 19104-6389, USA
}
\summary{%
NLTK, the Natural Language Toolkit, is a suite of program modules,
tutorials and problem sets, providing ready-to-use CL courseware.
NLTK covers symbolic and statistical
NLP, and is interfaced to annotated corpora.
Students augment and replace existing components, learning
structured programming by example, and manipulating sophisticated
models from the outset.}
\paperid{Pxxxx}
\keywords{courseware, Python, corpora, keyword4?, keyword5?}
\contact{Edward Loper}
\conference{this paper has not been submitted to any other conferences}
\date{\today}

@@

\section{Basic modules}

\subsection{Tree}

\subsection{Tagger}

\subsection{Probability}

\section{Advanced modules}

\subsection{ChunkParser}

\subsection{ChartParser}

\subsection{FSA}

\subsection{Classifier}

@@

How we used NLTK at Penn (including the competition).

Strengths and weaknesses

@@

>>>>

\documentclass[11pt]{article}
\usepackage{acl02sub,url,alltt,epsfig}
\title{NLTK: The Natural Language Toolkit}
\author{
Edward Loper and Steven Bird\\
Department of Computer and Information Science and Linguistic Data Consortium\\
University of Pennsylvania, Philadelphia, PA 19104-6389, USA
}
\summary{%
NLTK, the Natural Language Toolkit, is a suite of program modules,
tutorials and problem sets, providing ready-to-use CL courseware.
NLTK covers symbolic and statistical
NLP, and is interfaced to annotated corpora.
Students augment and replace existing components, learning
structured programming by example, and manipulating sophisticated
models from the outset.}
\paperid{Pxxxx}
\keywords{teaching, courseware, Python, corpora}
\contact{Edward Loper}
\conference{this paper has not been submitted to any other conferences}
\date{\today}

\newenvironment{sv}{\scriptsize\begin{alltt}}{\end{alltt}\normalsize}

@@

\section{Uses of NLTK}

\subsection{Assigments and projects}

NLTK can be used for simple and advanced projects, as well as
everything in between.  In the simplest assignments, students
tweak an existing module to alter its behavior.  The variety of
existing modules makes it easy to create simple assignments.
A more challenging task is to develop a new module.  Here, NLTK provides
some useful starting points: predefined interfaces and data structures, and
existing modules that implement the same interface.  Advanced projects
typically involve the development of entirely new functionality for a
previously unsupported NLP task, or the development of a complete system
out of existing and new modules.

Here we illustrate the use of NLTK in an assignment of moderate
difficulty, for the case of chunk parsing.  This is a kind of parsing
where we only identify the main constituents of a phrase, a technique
with diverse applications.  The NLTK \texttt{chunk} module offers
a variety of rule formats.  For example,
\texttt{ChunkRule('<NN.*>')} builds chunks over consecutive noun tags;
\texttt{ChinkRule('<VB.>')} excises verbs from existing chunks;
\texttt{SplitRule('<NN>', '<DT>')} splits into two pieces any existing
chunk that contains a noun followed by a determiner; and
\texttt{MergeRule('<JJ>', '<JJ>')} combines two adjacent chunks
where the first chunk ends and the second chunk starts with \texttt{JJ}.
The goal of the assignment is to come up with a cascade of these rules,
and to create a set of chunks that are as similar as possible to the test set.

The chunking tutorial provides all the necessary code except for the
chunk rules.  The provided code is responsible for loading the
chunked, tagged text using an existing tokenizer, creating an
unchunked version of the text, applying the chunk rules to the
unchunked text, and scoring the result.  Students focus on the
NLP task only -- providing a ruleset with the best coverage.

In the remainder of this section we reproduce some of the cascades
created by the students.  The first example illustrates a combination
of several rule types:

\begin{sv}
cascade = [
  ChunkRule('<DT><NN.*><VB.><NN.*>'),
  ChunkRule('<DT><VB.><NN.*>'),
  ChunkRule('<.*>'),
  UnChunkRule('<IN|VB.*|CC|MD|RB.*>'),
  UnChunkRule("<,|\.|``|''>"),
  MergeRule('<NN.*|DT|JJ.*|CD>', '<NN.*|DT|JJ.*|CD>'),
  SplitRule('<NN.*>', '<DT|JJ>')
]
\end{sv}

This next example illustrates a brute-force approach.  All tags most
commonly found inside of chunks are chunked, while other tags are excluded.

\begin{sv}
cascade = [
  ChunkRule('<\$|CD|DT|EX|PDT
             |PRP.*|WP.*|\#|FW
             |JJ.*|NN.*|POS|RBS|WDT>*')
]
\end{sv}

The third example involves putting everything in a single chunk,
then excising elements that do not belong.

\begin{sv}
cascade = [
  ChunkRule('<.*>+')
  ChinkRule('<VB.*|IN|CC|R.*|MD|WRB|TO|.|,>+')
]
\end{sv}

The main program which surrounds the student's work is moderately
simple as well.  The line replaced by students is underlined.

\begin{sv}
from nltk.token import *
from nltk.chunkparser import *
from nltk.rechunkparser import *
import os

path = "/cis530/data/wsj/tagged/"

sent_tok = RETokenizer('(=======+)?\\n\\n+', 0)
sentences = []
files = os.listdir(path)
for file in files:
    print 'Loading %s...' % file
    text = open(path+file).read()
    sentences += sent_tok.tokenize(text, source=file, unit='s')

chunkscore = ChunkScore()

\underline{cascade = [ ??? ]}
chunkparser = REChunkParser(cascade)

ctt = ChunkedTaggedTokenizer()
for s in sentences:
    correct = ctt.tokenize(s.type(), source=s.loc())
    unchunked = unchunk(correct)
    guess = chunkparser.parse(unchunked)
    chunkscore.score(correct, guess)

print chunkscore
\end{sv}

The score for each project is the precision, recall and F-measure.
Figure~\ref{contest} shows the results; students with the best two scores
were presented with prizes.

\begin{figure}
\centerline{\epsfig{figure=contest.ps,width=\linewidth}}
\caption{Precision/Recall Graph for Chunking Competition}\label{contest}
\vspace*{2ex}\hrule
\end{figure}

\subsection{Class demonstrations}

[chart parsing example, including screenshot]

@@

About CIS-530: mixture of CS and linguistics students.
Approx 40 hours of classes.  All class materials (slides,
assignments) are posted on the course website
\url{http://www.cis.upenn.edu/~cis530/}.

The Language: Python was new to 95\% of students.  Some were initially
bothered about its block structuring, which is sensitive to white
space.  All quickly learnt the language, and reported that any initial
reservations were unfounded.  Many students said that they enjoyed
writing Python code.

CD-ROM: put NLTK and data on CD-ROM and gave it out at the start of class.

The toolkit: students liked being able to do interesting projects from
the outset, liked all the documentation, liked being able to run
everything on their machine at home, ...

Problems: available corpus data, ...

Other strengths / weaknesses

@@

NOTES: Many CL textbooks provide good exercises at the end of each
chapter.  This isn't enough for many situations.  E.g. students spend
lots of time writing interface code (e.g. to process a corpus).
E.g. teachers spend lots of time creating do-able assignments...
